First Order Alternation

### 2019-02-05 quantifiers/body.tex
ok

\begin{definition}\label{def:unfolding}
  An \emph{unfolding} of an automaton $\A =
  \tuple{\Sigma,X,Q,\iota,F,\Delta}$ is a finite partial mapping $U :
  \Sigma^* \rightharpoonup_{\mathit{fin}} \posforms(Q,\emptyset)$,
  whose domain $\dom(U)$ is a finite prefix-closed complete set, such
  that $U(\epsilon) = \iota$, and for each sequence $\alpha a \in
  \dom(U)$, such that $\alpha \in \Sigma^*$ and $a \in \Sigma$:
  \[\begin{array}{c}
  \stamp{U(\alpha)}{0} \wedge
  \bigwedge_{q(\vec{y}) \arrow{a(X)}{} \psi} \forall y_1 \ldots
  \forall y_{\#q} ~.~ \stamp{q}{0}(\vec{y}) \rightarrow
  \stamp{\psi}{1} \models \stamp{U(\alpha a)}{1}
  \end{array}\]  
  Moreover, $U$ is \emph{safe} if for each $\alpha \in \dom(U)$, the
  formula $U(\alpha) \wedge \bigwedge_{q \in Q \setminus F} \forall
  y_1 \ldots \forall_{y_{\#(q)}} ~.~ q(\vec{y}) \rightarrow \bot$ is
  unsatisfiable.
\end{definition}

@@

\begin{definition}\label{def:generalized-lyndon-interpolant}
  Given a non-empty sequence of input events $\alpha = a_1 \ldots a_n
  \in \Sigma^*$, a \emph{generalized Lyndon interpolant (GLI)} is a
  sequence $(I_0,\ldots,I_n)$ of formulae such that, for all $k \in
  [n-1]$, the following hold: \begin{compactenum}
    \item\label{it1:generalized-lyndon-interpolant} $\pset{-}{I_k} =
      \emptyset$,
      %
    \item\label{it2:generalized-lyndon-interpolant} $\stamp{\iota}{0}
      \models I_0$ and \(I_k \wedge \Big(\bigwedge_{q(\vec{y})
        \arrow{a_i(X)}{} \psi \in \Delta} \forall y_1 \ldots \forall
      y_{\#(q)} ~.~ \stamp{q}{k}(\vec{y}) \rightarrow
      \stamp{\psi}{k+1}\Big) \models I_{k+1}\),
      %
    \item\label{it3:generalized-lyndon-interpolant} $I_n \wedge
      \bigwedge_{q \in Q \setminus F} \forall y_1 \ldots \forall
      y_{\#(q)} ~.~ q(\vec{y})$ is unsatisfiable.
  \end{compactenum}
  Moreover, the GLI is \emph{local} if and only if $\voc{I_k}
  \subseteq \stamp{Q}{k}$, for all $k \in [n]$.
\end{definition}
The following proposition states the existence of local GLI for the
theories in which Lyndon's Interpolation Theorem holds.

@@

\begin{theorem}[\cite{BorgerGraedelGurevich97}]\label{thm:skolem}
  Given $Q_1 x_1 \ldots Q_n x_n ~.~ \phi$ a first order sentence,
  where $Q_1, \ldots, Q_n \in \set{\exists,\forall}$ and $\phi$ is
  quantifier-free, let $\eta_i \isdef f_i(y_1,\ldots,y_{k_i})$ if $Q_i
  = \forall$ and $\eta_i \isdef x_i$ if $Q_i = \exists$, where $f_i$
  is a fresh function symbol and $\set{y_1, \ldots, y_{k_i}} =
  \set{x_j \mid j < i,~ Q_j = \exists}$. Then the entailment \(Q_1 x_1
  \ldots Q_n x_n ~.~ \phi \models \phi[\eta_1/x_1,\ldots,\eta_n/x_n]\)
  holds.
\end{theorem}
\proof{See \cite[Theorem 2.1.8]{BorgerGraedelGurevich97} and
  \cite[Lemma 2.1.9]{BorgerGraedelGurevich97}. \qed}

@@

\vspace*{-0.5\baselineskip}
\section{Experimental Results}
\vspace*{-0.5\baselineskip}

@@

@@

\begin{table}[htb]
%\vspace*{-\baselineskip}
\begin{center}
{\fontsize{7}{8}\selectfont
\begin{tabular}{||l|c|c|c|c|c||}
\hline
Example & $\len{\A}$ (bytes) & $L(\A)=\emptyset$ ? & Nodes Expanded & Nodes Visited & Time (ms) \\
\hline
incdec.pa & 499 & no & 21 & 17 & 779 \\
\hline
localdec.pa & 678 & no & 49 & 35 & 1814 \\
\hline
ticket.pa & 4250 & no & 229 & 91 & 9543 \\
\hline
count\_thread0.pa & 9767 & no & 154 & 128 & 8553 \\
\hline
count\_thread1.pa & 10925 & no & 766 & 692 & 76771 \\
\hline
local0.pa & 10595 & no & 73 & 27 & 1431 \\
\hline
local1.pa & 11385 & no & 1135 & 858 & 101042 \\
\hline
array\_rotation.ada & 1834 & yes & 9 & 8 & 1543 \\
\hline
array\_simple.ada & 3440 & yes & 11 & 10 & 6787 \\
\hline
array\_shift.ada & 874 & yes & 6 & 5 & 413 \\
\hline
abp.ada & 6909 & no & 52 & 47 & 4788 \\
\hline
train.ada & 1823 & yes & 68 & 67 & 7319 \\
\hline
hw1.ada & 322 & Solver Error & / & / & / \\
\hline
hw2.ada & 674 & yes & 20 & 22 & 4974 \\
\hline
rr-crossing.foada & 1780 & yes & 67 & 67 & 7574 \\
% \hline
% running2.foada & 1217 & yes & 13 & 13 & 788 \\
\hline
train-simple1.foada & 5421 & yes & 43 & 44 & 2893 \\
\hline
train-simple2.foada & 10177 & yes & 111 & 113 & 8386 \\
\hline
train-simple3.foada & 15961 & yes & 196 & 200 & 15041 \\
% \hline
% fischer-mutex1.foada & 1566 & no & 31 & 31 & 1552 \\
\hline
fischer-mutex2.foada & 3000 & yes & 23 & 23 & 808 \\
\hline
fischer-mutex3.foada & 4452 & yes & 33 & 33 & 1154 \\
\hline
\end{tabular}
}
\caption{Experiments with First Order Alternating Automata}
\label{tab:experiments}
\end{center}
\vspace*{-2\baselineskip}
\end{table}

\section{Conclusions}

>>>>

\begin{definition}\label{def:unfolding}
  An \emph{unfolding} of an automaton $\A =
  \tuple{\Sigma,X,Q,\iota,F,\Delta}$ is a finite partial mapping $U :
  \Sigma^* \rightharpoonup_{\mathit{fin}} \posforms(Q,\emptyset)$,
  whose domain $\dom(U)$ is a finite prefix-closed complete set, such
  that $U(\epsilon) = \iota$, and for each sequence $\alpha a \in
  \dom(U)$, such that $\alpha \in \Sigma^*$ and $a \in \Sigma$:
  \[\begin{array}{c}
  \stamp{U(\alpha)}{0} \wedge
  \bigwedge_{q(\vec{y}) \arrow{a(X)}{} \psi} \forall y_1 \ldots
  \forall y_{\#q} ~.~ \stamp{q}{0}(\vec{y}) \rightarrow
  \stamp{\psi}{1} \models \stamp{U(\alpha a)}{1}
  \end{array}\]  
  $U$ is \emph{safe} if for each $\alpha \in \dom(U)$, the formula
  $U(\alpha) \wedge \bigwedge_{q \in Q \setminus F} \forall y_1 \ldots
  \forall_{y_{\#(q)}} ~.~ q(\vec{y}) \rightarrow \bot$ is
  unsatisfiable.
\end{definition}

@@

\begin{definition}\label{def:generalized-lyndon-interpolant}
  Given a non-empty sequence of input events $\alpha = a_1 \ldots a_n
  \in \Sigma^*$, a \emph{generalized Lyndon interpolant (GLI)} is a
  sequence $(I_0,\ldots,I_n)$ of formulae such that, for all $k \in
  [n-1]$, the following hold: \begin{compactenum}
    \item\label{it1:generalized-lyndon-interpolant} $\pset{-}{I_k} =
      \emptyset$,
      %
    \item\label{it2:generalized-lyndon-interpolant} $\stamp{\iota}{0}
      \models I_0$ and \(I_k \wedge \Big(\bigwedge_{q(\vec{y})
        \arrow{a_i(X)}{} \psi \in \Delta} \forall y_1 \ldots \forall
      y_{\#(q)} ~.~ \stamp{q}{k}(\vec{y}) \rightarrow
      \stamp{\psi}{k+1}\Big) \models I_{k+1}\),
      %
    \item\label{it3:generalized-lyndon-interpolant} $I_n \wedge
      \bigwedge_{q \in Q \setminus F} \forall y_1 \ldots \forall
      y_{\#(q)} ~.~ q(\vec{y}) \rightarrow \bot$ is unsatisfiable.
  \end{compactenum}
  Moreover, the GLI is \emph{local} if and only if $\voc{I_k}
  \subseteq \stamp{Q}{k}$, for all $k \in [n]$.
\end{definition}
The following proposition states the existence of local GLI for the
theories in which Lyndon's Interpolation Theorem holds.

@@

\begin{theorem}[\cite{BorgerGraedelGurevich97}]\label{thm:skolem}
  Given $Q_1 x_1 \ldots Q_n x_n ~.~ \phi$ a first order sentence,
  where $Q_1, \ldots, Q_n \in \set{\exists,\forall}$ and $\phi$ is
  quantifier-free, let $\eta_i \isdef f_i(y_1,\ldots,y_{k_i})$ if $Q_i
  = \forall$ and $\eta_i \isdef x_i$ if $Q_i = \exists$, where $f_i$
  is a fresh function symbol and $\set{y_1, \ldots, y_{k_i}} =
  \set{x_j \mid j < i,~ Q_j = \exists}$. Then the entailment \(Q_1 x_1
  \ldots Q_n x_n ~.~ \phi \models \phi[\eta_1/x_1,\ldots,\eta_n/x_n]\)
  holds.
\end{theorem}
\paragraph{Proof.} See \cite[Theorem 2.1.8]{BorgerGraedelGurevich97} and
  \cite[Lemma 2.1.9]{BorgerGraedelGurevich97}. \qed

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\enlargethispage{6mm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experimental Results}

@@

\begin{table}[htb]
\vspace*{-\baselineskip}
\begin{center}
{\fontsize{7}{8}\selectfont
\begin{tabular}{||l|c|c|c|c|c||}
\hline
Example & $\len{\A}$ (bytes) & $L(\A)=\emptyset$ ? & Nodes Expanded & Nodes Visited & Time (ms) \\
\hline
incdec.pa & 499 & no & 21 & 17 & 779 \\
\hline
localdec.pa & 678 & no & 49 & 35 & 1814 \\
\hline
ticket.pa & 4250 & no & 229 & 91 & 9543 \\
\hline
count\_thread0.pa & 9767 & no & 154 & 128 & 8553 \\
\hline
count\_thread1.pa & 10925 & no & 766 & 692 & 76771 \\
\hline
local0.pa & 10595 & no & 73 & 27 & 1431 \\
\hline
local1.pa & 11385 & no & 1135 & 858 & 101042 \\
\hline
array\_rotation.ada & 1834 & yes & 9 & 8 & 1543 \\
\hline
array\_simple.ada & 3440 & yes & 11 & 10 & 6787 \\
\hline
array\_shift.ada & 874 & yes & 6 & 5 & 413 \\
\hline
abp.ada & 6909 & no & 52 & 47 & 4788 \\
\hline
train.ada & 1823 & yes & 68 & 67 & 7319 \\
\hline
hw1.ada & 322 & Solver Error & / & / & / \\
\hline
hw2.ada & 674 & yes & 20 & 22 & 4974 \\
\hline
rr-crossing.foada & 1780 & yes & 67 & 67 & 7574 \\
% \hline
% running2.foada & 1217 & yes & 13 & 13 & 788 \\
\hline
train-simple1.foada & 5421 & yes & 43 & 44 & 2893 \\
\hline
train-simple2.foada & 10177 & yes & 111 & 113 & 8386 \\
\hline
train-simple3.foada & 15961 & yes & 196 & 200 & 15041 \\
% \hline
% fischer-mutex1.foada & 1566 & no & 31 & 31 & 1552 \\
\hline
fischer-mutex2.foada & 3000 & yes & 23 & 23 & 808 \\
\hline
fischer-mutex3.foada & 4452 & yes & 33 & 33 & 1154 \\
\hline
\end{tabular}
}
\caption{Experiments with First Order Alternating Automata}
\label{tab:experiments}
\end{center}
\vspace*{-3\baselineskip}
\end{table}

@@

\vspace*{-\baselineskip}
\section{Conclusions}
